\documentclass[99-notes-packed.tex]{subfiles}

\begin{document}
\paragraph*{Concurrency Semantics} Define the following semantics of interest wrt. concurrency theory: 
\begin{enumerate}
    \item {
        \textbf{Interleaving Semantics}: concurrent actions $a$, $b$ occur in one of the following orders: 
        \begin{itemize}
            \item $a; b$
            \item $b; a$
        \end{itemize}
    }
    \item {
        \textbf{Step Semantics}: concurrent actions $a$, $b$ occur in one of the following orders: 
        \begin{itemize}
            \item $a; b$
            \item $b; a$
            \item $a || b$ (in parallel)
        \end{itemize}
    }
    \item {
        \textbf{Interval Semantics}: concurrent actions $a$, $b$ occur in continuous time, such that $a$, $b$ may occur in parallel for a subset of total runtime. 
    }
    \item {
        \textbf{Partial-order (aka. Causal) Semantics}: concurrent actions $a$, $b$ not only can occur in parallel for some continuous time interval, but can also intersperse as unspecified segments (e.g., OS scheduling). 

        Nevertheless, causal relationships between $a, b, \dots$ are preserved -- $a$ calling e.g. \texttt{fork()} will posit a partial ordering before the spawned task $b$, though the exact concurrency behaviors leave much leeway to the OS scheduler.
    }
\end{enumerate}

\begin{definition}[Pomset]
    A \textbf{pomset (partial-ordered multiset)} defines a $(E, <, l)$-tuple where:
    \begin{itemize}
        \item $E$ a set of ``events'' -- corresponding to each occurrence of action. 
        \item $<$ a partial-order of E such that $e_i$ happens before $e_j$, or incomparable, etc.
        \item $l: E \rightarrow A$ a mapping between events to their actions. 
    \end{itemize}

    A normal trace thus becomes a totally ordered multiset of actions, compared to a pomset representation. 
\end{definition}

\paragraph*{Petri Net} captures the dynamism within parallel systems. It defines a $(S, T, F, I)$-tuple where: 
\begin{itemize}
    \item $S, T$ define \underline{places} and actions grouped in bipartite form. 
    \item $F \subseteq (S \times T) \cup (T \times S)$ set of transitions.
    \item $I: S \rightarrow \mathbb{N}$ (initial marking) defines the initial state of control, via allocating tokens to initial states. Subsequent states of control is referred to as \textbf{marking} in general.
\end{itemize}

\begin{definition}[Control \& Tokens]
    Petri nets encode control at runtime. A \textbf{control} simply refers to the state at which the system is currently at. It is represented as a assignment of \textbf{tokens} to each place in net i.e. a \textbf{marking}. 
\end{definition}

\paragraph*{Progress, Justness, Fairness} We define several properties on concurrent systems to make claim they display good behaviors. 

\begin{definition}[Safety]
    A \textbf{safety property} defines that a ``bad'' predicate $\phi$ would never hold -- e.g., $P \models G(\neg \phi)$. 
\end{definition}

\begin{definition}[Liveness]
    A \textbf{liveness property} defines that a ``good'' predicate $\phi$ would eventually hold -- e.g., $P \models F(\phi)$. 
\end{definition}

It's easy to see that safety and liveness properties are duo/convertible to each other -- $G(\neg \phi) \iff \neg F(\phi)$, after all. Hence we can speak of them in common contexts.

Whether safety/liveness properties hold in a system depends often on whether we make appropriate progress and fairness assumptions (in increasing hierarchy): 

\begin{background}[Completeness Criteria]
    LTL is a \textbf{linear-time logic} that specifies \textbf{linear-time properties} on paths in a Kripke structure. For example, $F(\phi)$ really means that, within each/a path, a state marked with proposition $\phi$ eventually occurs within the path string. 

    We hence expand the satisfaction idea such that a process $P$ satisfies a LTL property $\varphi$ under a \textbf{completeness criterion} $C$: 
    \begin{equation*}
        P \models^{C} \varphi \iff \forall\ \mathrm{path}\ \pi \in P: \pi \models C \implies \pi \models \varphi
    \end{equation*}

    For example, $C$ might refer to the assumption that a path is infinite -- \textbf{deadlock-free}, as a finite path occurs only if we reach a state without outgoing transitions. This is often the \textbf{default completeness criterion} when unspecified -- in which case we simply use $\models$ without superscript. 

    A completeness criterion $D$ is \textbf{stronger} than $C$ if: 
    \begin{equation*}
        \{\pi\ |\ \pi \in \mathsf{path}(P), \pi \models D\} \subset \{\pi\ |\ \pi \in \mathsf{path}(P), \pi \models C\}
    \end{equation*}
\end{background}

We hence define the following completeness criteria in increasing strength (or reverse implication chain): 

\begin{definition}[Progress]
    In a \textbf{closed system} (i.e., without external influence of transition decisions), the \textbf{progress} assumption assumes that we will not stay forever in a state with an outgoing transition -- i.e., a finite path is complete iff it ends in a state without outgoing transitions.

    To show this is weaker than weak fairness, simply show that task $T$ cannot occur whatsoever in last state of a non-progress path (because it is enabled, but trivially not occurred -- because of non-progression past that state). 
\end{definition}

There is also \textbf{justness} but this was not covered in lecture.

\begin{definition}[Weak Fairness]
    Define a \textbf{task} to be a set of transitions in a Kripke structure (you can think of it as a subprocedure in a program). Append the Kripke structure $(S, \rightarrowtriangle, I)$ with a novel structure $\tau$: 
    \begin{equation*}
        \tau \coloneqq \textrm{Collection of tasks}
    \end{equation*}

    A task $T$ is \textbf{enabled} in state $s \in S$ if there exists an outgoing transition from $s$ that is also in $T$. $T$ is \textbf{perpetually enabled} on path $\pi$ if it is enabled in every state of $\pi$ (i.e., every state in $\pi$ contains outgoing transition in $T$). 
    
    Orthogonally, if a $T$-transition exists \underline{in} $\pi$ then $T$ \textbf{occurs} in $\pi$.

    A path $\pi$ is hence \textbf{weakly fair} if it satisfies the following LTL formula:
    \begin{alignat*}{2}
        \mathsf{WF}(T) 
        &\coloneqq &&\ G[G(\mathsf{enabled}(T)) \implies F(\mathsf{occurs(T)})] \\ 
        &\Leftrightarrow &&\ F(G(\mathsf{enabled}(T))) \implies G(F(\mathsf{occurs}(T)))
    \end{alignat*}

    \begin{quote}
        Read it in this way: if, after some $s \in \pi$, part of T can always happen for $s^{'} \in \pi^{'}$ (i.e., perpetually enabled), then at least some part of T \underline{will} happen in $\pi^{'}$.
    \end{quote}

    where predicates $\mathsf{enabled}$, $\mathsf{occurs}$ follow above definition -- a state $s$ is marked with $\mathsf{enabled}(T)$ iff $T$ is enabled at state $s$ i.e. exists outgoing transition from $s$ that is also in $T$.

    Hence, define $\models^{WF}$ as: 
    \begin{equation*}
        P \models^{WF} \varphi \iff P \models [\wedge_{T \in \tau} {(\mathsf{WF}(T))}] \implies \varphi
    \end{equation*}
\end{definition}

\begin{definition}[Strong Fairness]
    A task $T$ is \textbf{relentlessly enabled} on $\pi$ if for every suffix of $\pi$ there exists a state $s^{'}$ where $T$ is enabled (note the difference between perpetually enabled). 

    A path $\pi$ is hence \textbf{strongly fair} if it satisfies the following LTL formula: 
    \begin{alignat*}{2}
        \mathsf{SF}(T) &\coloneqq\ && G[GF(\mathsf{enabled}(T)) \implies F(\mathsf{occurs}(T))] \\
        &\Leftrightarrow\ && GF(\mathsf{enabled}(T)) \implies GF(\mathsf{occurs}(T))
    \end{alignat*}

    \begin{quote}
        In other words: if, after some $s \in \pi$, part of T will always be available occasionally for some $s^{'} \in \pi^{'}$ (i.e., relentlessly enabled), then some part of T \underline{will} happen in $\pi^{'}$. 
    \end{quote}
    
    Likewise, define $\models^{SF}$ in mirror case of $\models^{WF}$. 
\end{definition}

\end{document}